Reconstruction of Large Scale Structure by Nonlinear \tcr{Lagrangian} Displacement Fields

### 2016-10-04 ms_disp1/ms_disp1.tex
oct 4

\addtolength{\hoffset}{-0.525cm}
\addtolength{\textwidth}{1.05cm}
\title{Reconstruction of Large Scale Structure by Nonlinear \tcr{Lagrangian} Displacement Fields - I}

@@

Standard BAO reconstruction algorithms smooth the 
nonlinear density field on linear scale ($\sim$10 Mpc$/h$) and reverse the large 
scale bulk flows by a negative Zel'dovich linear displacement
\citep{2007ApJ...664..675E,2009PhRvD..80l3501N,2009PhRvD..79f3523P}. Here we propose a 
\tcr{new reconstruction method} that uses the nonlinear displacement field to recover the 
primordial density field. In the linear Lagrangian perturbation theory, the 
negative divergence of the the displacement field
$\bs\Psi(\bs q)$ respect to Lagrangian coordinates $\bs q$
gives the linear density field \tcr{\citep{2010PhDT.........4J}}.
Of course, the full displacement field $\bs\Psi(\bs q)$
is non-observable, as it requires the initial distributions of matter, however there are 
many techniques to estimate the nonlinear displacement field from a final 
distribution of matter. For example, when a homogeneous initial matter distribution 
is assumed, there is a unique solution of curl-less displacement field to relate 
the initial and final distributions without shell-crossing. This solution can be 
solved by a metric transformation equation \citep{1995ApJS..100..269P,1998ApJS..115...19P}.
\cite{1989ApJ...344L..53P,2000ApJ...544...21G}
\tcr{[methods of Peebles, david Goldberg? least action ]}
In 1-dimensional (1D) case, the exact solution of \cite{1995ApJS..100..269P}
simplifies to an ordering of matter elements by Eulerian coordinates.
Zhu {\it et al.} \cite{2016arXiv160907041Z} apply this algorithm to the result
of an 1D simulation \citep{2016JCAP...01..043M} and obtain an estimated
displacement field $s(q)$, and find that \tcr{this new method} well recovers the linear
information and reconstructs BAO peak in correlation function. 
In 3D case, there are various techniques to obtain $\bs s(\bs q)$.
However one needs to carefully consider effects of curl, shell-crossing in 3D case.

Before these steps, we need to quantify the amount of linear information
that can be recovered from the full nonlinear displacement field $\bs\Psi(\bs q)$,
and further estimations $\bs s(\bs q)$ can be compared with this result.
In this paper we run a LSS $N$-body simulation and track the motion of
particles to obtain $\bs\Psi(\bs q)$. According to this we reconstruct the linear density
field and compare to the primordial density field of the initial
conditions. We describe the simulation and reconstruction algorithm in section
\ref{sec.method}, and we show the results in section \ref{sec.results}.
Discussion and conclusion are in section \ref{sec.discussion}.

@@

\subsection{Simulation}\label{ss.sim}
We use the open source cosmological simulation code {\tt CUBE}%\footnote{github link of cafcube}
\citep{cafcube}.
Cosmological parameters are set as
$\Omega_m=0.27$, $\Omega_\Lambda=0.73$, $h_0=0.68$, $n_s=0.96$ and $\sigma_8=0.83$.
Initial conditions are generated at redshift $z=50$ 
using Zel'dovich approximation, and using a CLASS transfer function.
$N_p=512^3$ $N$-body particles are evolved via 
their mutual gravitational interactions to $z=0$, in a periodic box with $L=300$ 
Mpc$/h$ per side. The code is set to use standard a particle-mesh (PM) algorithm 
\cite{1988csup.book.....H} on a two-level mesh grids
(details see \cite{2013MNRAS.436..540H}) and cloud-in-cell
(CIC) is used in particle interpolations in force 
calculation and obtaining the density field $\rho({\bs x})$ in Eulerian coordinates 
${\bs x}$ at late stages. We use density contrast $\delta\equiv\rho/\lb\rho\rb-1$ 
to describe the density fluctuations. The primordial linear density field $
\delta_L$ is given by the initial stage and scaled to $z=0$ by the linear growth 
factor.

@@

\section{Results}\label{sec.results}
Fig.\ref{fig.2} shows the cross correlation functions,
damping factors $W^2(k)$ for the optimal filtered 
nonlinear and reconstructed density fields.
We can see that the cross correlation $r_{LR}$ is recovered to
nearly an order of magnitude smaller scales compared to $r_{LN}$.
We also fit the Gaussian BAO damping model $
{\mathcal D}(k)=\exp(-k^2\Sigma^2/2)$ and give $\Sigma=1.8$ Mpc$/h$ and $\Sigma=10$ 
Mpc$/h$ for nonlinear and reconstructed fields.
We repeat the analysis with various box sizes,
and the results are consistent.

In Fig.\ref{fig.3} and Fig.\ref{fig.4} we show the power spectrum
decomposition by equation (\ref{eq.power}) for both $\delta_N$ 
(with subscripts ``$_R$'' replaced by ``$_N$'') and $\delta_R$.
In Fig.\ref{fig.3} we can see that, on scales $k>0.2\ h/$Mpc,
$P_{NN}$ term dominates the power spectrum due to a low correlation
between $\delta_L$ and $delta_N$. The optimal filtered power spectrum
damps quickly on small scales. In comparison, in Fig.\ref{fig.4},
$P_{NN}$ starts to dominate on scales $k>1\ h/$Mpc. As a result, the
optimal filtered reconstructed power spectrum
(by equation (\ref{eq.opt})) is much higher.

\section{Discussion and conclusion}\label{sec.discussion}
We extract the actual displacement field of matter elements in cosmological $N$-body
simulations, and use this displacement field in Lagrangian coordinates to reconstruct
the primordial linear perturbations. The result shows a prominent improvement from
$r_{LN}$ to $r_{LR}$ in Fig.\ref{fig.2} -- recovering the lost linear information on
nearly an order of magnitude smaller scales.
This is achieved by implementing differential movement information
of matter elements on Lagrangian coordinates, rather than on
Eulerian coordinates. This result illustrates the feasibility
of using estimated displacement field $\bs s(\bs q)$ to reconstruct primordial linear
density field. A straightforward example of a estimation of $\bs s(\bs q)$
is given by \cite{1995ApJS..100..269P,1998ApJS..115...19P}.
More complicated algorithms can be found in \tcr{[cite]}, where
shell-crossing effects are also considered. In reality, one needs
to consider all aspects including vorticity, shell-crossing, bias, noise
and data complexities. The impact of these factors can be quantitatively
compared with the impact of different estimation methods, and with
the exact solution by $N$-body simulations.

@@

\section*{Acknowledgements}
HRY acknowledges General Financial Grant from the China Postdoctoal Science Foundation No.2015M570884.

>>>>

\addtolength{\hoffset}{-0.525cm}
\addtolength{\textwidth}{1.05cm}
\title{Reconstruction of Large Scale Structure by Nonlinear \tcr{Lagrangian} Displacement Fields}

@@

Standard BAO reconstruction algorithms smooth the 
nonlinear density field on linear scale ($\sim$10 Mpc$/h$) and reverse the large 
scale bulk flows by a negative Zel'dovich linear displacement
\citep{2007ApJ...664..675E,2009PhRvD..80l3501N,2009PhRvD..79f3523P}. Here we propose a 
\tcr{new reconstruction method} that uses the nonlinear displacement field to recover the 
primordial density field. In the linear Lagrangian perturbation theory, the 
negative divergence of the the displacement field
$\bs\Psi(\bs q)$ respect to Lagrangian coordinates $\bs q$
gives the linear density field \tcr{\citep{2010PhDT.........4J}}.
Of course, the full displacement field $\bs\Psi(\bs q)$
is non-observable, as it requires the initial distributions of matter, however there are 
many techniques to estimate the nonlinear displacement field from a final 
distribution of matter. For example, when a homogeneous initial matter distribution 
is assumed, there is a unique solution of curl-less displacement field to relate 
the initial and final distributions without shell-crossing. This solution can be 
solved by a metric transformation equation \citep{1995ApJS..100..269P,1998ApJS..115...19P}.
In 1-dimensional (1D) case, the exact solution of \cite{1995ApJS..100..269P}
simplifies to an ordering of matter elements by Eulerian coordinates.
Zhu {\it et al.} \cite{2016arXiv160907041Z} apply this algorithm to the result
of an 1D simulation \citep{2016JCAP...01..043M} and obtain an estimated
displacement field $\tilde\Psi(q)$, and find that \tcr{this new method} well recovers the linear
information and reconstructs BAO peak in correlation function. 
In 3D case, there are various techniques to obtain $\tilde{\bs \Psi}(\bs q)$.
However one needs to carefully consider effects of curl, shell-crossing in 3D case.

Before these steps, we need to quantify the amount of linear information
that can be recovered from the full nonlinear displacement field $\bs\Psi(\bs q)$,
and further estimations $\tilde{\bs \Psi}(\bs q)$ can be compared with this result.
In this paper we run a LSS $N$-body simulation and track the motion of
particles to obtain $\bs\Psi(\bs q)$. According to this we reconstruct the linear density
field and compare to the primordial density field of the initial
conditions. We describe the simulation and reconstruction algorithm in section
\ref{sec.method}, and we show the results in section \ref{sec.results}.
Discussion and conclusion are in section \ref{sec.discussion}.

@@

\subsection{Simulation}\label{ss.sim}
We use the open source cosmological simulation code {\tt CUBE} %\footnote{github link of cafcube}
\citep{cafcube}.
Cosmological parameters are set as
$\Omega_m=0.27$, $\Omega_\Lambda=0.73$, $h_0=0.68$, $n_s=0.96$ and $\sigma_8=0.83$.
Initial conditions are generated at redshift $z=50$ 
using Zel'dovich approximation, and using a CLASS transfer function.
$N_p=512^3$ $N$-body particles are evolved via 
their mutual gravitational interactions to $z=0$, in a periodic box with $L=300$ 
Mpc$/h$ per side. The code is set to use standard a particle-mesh (PM) algorithm 
\cite{1988csup.book.....H} on a two-level mesh grids
(details see \cite{2013MNRAS.436..540H}) and cloud-in-cell
(CIC) is used in particle interpolations in force 
calculation and obtaining the density field $\rho({\bs x})$ in Eulerian coordinates 
${\bs x}$ at late stages. We use density contrast $\delta\equiv\rho/\lb\rho\rb-1$ 
to describe the density fluctuations. The primordial linear density field $
\delta_L$ is given by the initial stage and scaled to $z=0$ by the linear growth 
factor.

@@

\section{Results}\label{sec.results}
Fig.\ref{fig.2} shows the cross correlation functions,
damping factors $W^2(k)$ for the optimal filtered 
nonlinear and reconstructed density fields.
We can see that the cross correlation $r_{LR}$ is recovered to
nearly an order of magnitude smaller scales compared to $r_{LN}$.
We also fit the Gaussian BAO damping model $
{\mathcal D}(k)=\exp(-k^2\Sigma^2/2)$ and give $\Sigma=1.8$ Mpc$/h$ and $\Sigma=10$ 
Mpc$/h$ for reconstructed and nonlinear fields.
We repeat the analysis with various box sizes,
and the results are consistent.

In Fig.\ref{fig.3} and Fig.\ref{fig.4} we show the power spectrum
decomposition by equation (\ref{eq.power}) for both $\delta_N$ 
(with subscripts ``$_R$'' replaced by ``$_N$'') and $\delta_R$.
In Fig.\ref{fig.3} we can see that, on scales $k>0.2\ h/$Mpc,
$P_{NN}$ term dominates the power spectrum due to a low correlation
between $\delta_L$ and $\delta_N$. The optimal filtered power spectrum
damps quickly on small scales. In comparison, in Fig.\ref{fig.4},
$P_{NN}$ starts to dominate on scales $k>1\ h/$Mpc. As a result, the
optimal filtered reconstructed power spectrum
(by equation (\ref{eq.opt})) is much higher.

\section{Discussion and conclusion}\label{sec.discussion}
We extract the actual displacement field of matter elements in cosmological $N$-body
simulations, and use this displacement field in Lagrangian coordinates to reconstruct
the primordial linear perturbations. The result shows a prominent improvement from
$r_{LN}$ to $r_{LR}$ in Fig.\ref{fig.2} -- recovering the lost linear information on
nearly an order of magnitude smaller scales.
This is achieved by implementing differential movement information
of matter elements on Lagrangian coordinates, rather than on
Eulerian coordinates. This result illustrates the feasibility
of using estimated displacement field $\tilde{\bs \Psi}(\bs q)$ to reconstruct primordial linear
density field. A straightforward example of a estimation of $\tilde{\bs \Psi}(\bs q)$
is given by \cite{1995ApJS..100..269P,1998ApJS..115...19P}.
More complicated algorithms can be found in \tcr{[cite]}, where
shell-crossing effects are also considered. In reality, one needs
to consider all aspects including vorticity, shell-crossing, bias, noise
and data complexities. The impact of these factors can be quantitatively
compared with the impact of different estimation methods, and with
the exact solution by $N$-body simulations.

\cite{1989ApJ...344L..53P,2000ApJ...544...21G}
\tcr{[methods of Peebles, david Goldberg? least action ]}

@@

\section*{Acknowledgements}
We thank Hong-Ming Zhu for helpful discussions.
HRY acknowledges General Financial Grant No.2015M570884 and Special Financial Grant No.2016T90009 from the China Postdoctoal Science Foundation.
